A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-26 paper/fascd.tex
implement figure reductions/removals, and some text edits; do more after finalizing results

\usepackage[kw]{pseudo}
\pseudoset{%
left-margin=15mm,%
topsep=5mm,%
label=\footnotesize\arabic*,%
idfont=\texttt,%
ctfont=\textsl,%
ct-left=\qquad\qquad(,%
ct-right=),%
}

@@

% Sets running headers as well as PDF title and authors
\headers{FAS for nonlinear variational inequalities}{E. Bueler and P. E. Farrell}

@@

\begin{abstract}
We present the \emph{full approximation scheme constraint decomposition} (FASCD) multilevel method for solving variational inequalities (VIs).  FASCD is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.  We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.  We demonstrate that FASCD V-cycles exhibit excellent mesh-independent convergence rates.  Furthermore, when a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD F-cycles (full multigrid) are optimal solvers.  The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.  Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
\end{abstract}

@@

The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.  These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.  In the multilevel case, this CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the algorithms of \cite{GraeserKornhuber2009,Tai2003} in the following directions:
\pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The method requires neither an objective function nor linearity of the residual.
\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
%\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
%\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (Algorithm \ref{alg:fascd-fcycle}) are proposed.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.  The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

Prior multilevel algorithms for VIs have various substantial limitations compared to the FASCD framework.  The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.  While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.  The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.  The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).  The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.  Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method, both simplifying the implementation and allowing the coarse grid corrections to update the estimate of the active set on the fine grid. Numerical evidence indicates that the FAS construction of coarse grid corrections is useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.
\pef{Insert comparison against the spiral problem once we have set rtol = 1e-12}
A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.

In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for corresponding unconstrained PDE problems.  The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.  Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.  Furthermore, these results appear to be insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that by \eqref{eq:fe:monotoneproperty} and \eqref{eq:fe:monotoneadditional},
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.  The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.

%REGENERATE using https://github.com/bueler/mg-glaciers:
%$ cd mg-glaciers/py/
%$ [modify decomposition() inside visualize.py to turn off axes and print \gamma for obstacle]
%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
\begin{figure}[ht]
\centering
\includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum injection: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
\label{fig:chiphilevels}
\end{figure}

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.  Correspondingly, there is another basic iteration that does not need the $\Pi_i$ operators, namely an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}).  Application \pr{icd-tele} below requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.

@@

While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.

@@

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.  Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the corrected finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\centering
\input{tikz/vcycle.tex}
\caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
\label{fig:fascdvcycle}
\end{figure}

@@

However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.  (In the VI context one must also truncate to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.

\begin{figure}[h]
\centering
\input{tikz/fcycle.tex}
\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray).  Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.  Solution on each level is by V-cycles. See Figure \ref{fig:fascdvcycle} for the other symbols.}
\label{fig:fcycle}
\end{figure}

@@

Table \ref{tab:results:classical} shows the resulting iteration counts.  As expected, RS Newton applied by itself as a solver shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.  These experiments over-solve, with \texttt{rtol} $=$ \texttt{atol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.  The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance.  For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution FIXME.  FIXME: asymp rate figure?

@@

\begin{example}  \label{ex:results:plap}
We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for nonlinear $p$-Laplacians with both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.  ``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty,0$, respectively; ellipticity is lost in the latter case.  Well-posedness and solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.

This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and the need for robust smoothers when the operator is nonlinear.  Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.  The exact solution is easily computed for any $p$; see Figure \ref{fig:results:plap} and the accompanying this paper.  For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems, an $O(m_j)$ smoother in 1D.  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.  Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.

%REGENERATE using elbueler/plapfigs branch of git@bitbucket.org:pefarrell/fascd.git, in examples/
% $ python3 plap1d.py -p 1.5 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d1p5.pdf
% $ python3 plap1d.py -p 4.0 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d4p0.pdf
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{fixfigs/plap1d1p5.pdf} \quad
\includegraphics[width=0.45\textwidth]{fixfigs/plap1d4p0.pdf}
\caption{Solutions $u\ge \psi$ of the $p$-Laplacian VI in Example \ref{ex:results:plap}, for $p=1.5$ (left) and $p=4.0$ (right).}
\label{fig:results:plap}
\end{figure}

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands (including imagemagick):
%   cd examples/
%   python3 plap1d.py -fascd_cycle_type multiplicative -mcoarse 4 -levs 3 -figures
%   [save the images into mcd-extended/paper/fixfigs/vcycle/]
%   for X in start2 down2 down1 up0 up1 up2 next2; do mogrify -trim $X.png $X.png; done
\begin{figure}[ht]
\centering
\input{tikz/imagesvcycle.tex}
\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.  Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines, obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots, and the zero function is dotted.  Vertical scale varies.}
\label{fig:imagesvcycle}
\end{figure}

Iteration results are shown in Table \ref{tab:results:fastplap1d} for the fast diffusion $p=1.5$ case.  We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).  F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.  If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').

Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.  V-cycles using the default settings, specifically three fixed RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.  Visualization shows an instability in which oscillations in the restricted residuals explode.  However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.  Even for the three-iteration smoother, a single F-cycle suffices at every resolution.

The $p=1.5$ and $p=4$ cases give different convergence rates in the $\|\cdot\|_\infty$ norm, namely at rates $O(h^{2.005})$ and $O(h^{1.420})$, respectively; compare \cite{ChoeLewis1991}.
\end{example}

>>>>

\usepackage[kw]{pseudo}
\pseudoset{%
left-margin=6mm,%
topsep=5mm,%
label=\footnotesize\arabic*,%
idfont=\texttt,%
ctfont=\textsl,%
ct-left=\qquad\qquad(,%
ct-right=),%
}

@@

% Sets running headers as well as PDF title and authors
\headers{FASCD for nonlinear variational inequalities}{E. Bueler and P. E. Farrell}

@@

\begin{abstract}
We present the \emph{full approximation scheme constraint decomposition} (FASCD) multilevel method for solving variational inequalities (VIs).  FASCD is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.  We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.  When a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit mesh-independent convergence rates and F-cycles (full multigrid) are optimal solvers.  The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
\end{abstract}

@@

The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.  In a finite element (FE) context, these methods converge for multilevel and domain decompositions.  The multilevel CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method enjoys the following advantages:

\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The method requires neither an objective function nor linearity of the residual.
\item Box constraints, i.e.~both upper and lower obstacles, are allowed.
\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (Algorithm \ref{alg:fascd-fcycle}) are proposed.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycles are an optimal solver for the examples in Section \ref{sec:results}; smoother effort is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

Prior multilevel algorithms for VIs have limitations compared to the FASCD framework.  The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.  While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.  The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.  The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).  The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.  Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method, both simplifying the implementation and allowing the coarse grid corrections to update the estimate of the active set on the fine grid. Numerical evidence indicates that our coarse grid corrections are useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.  On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).

A known limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.

In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for corresponding unconstrained PDE problems.  The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.  Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.  Furthermore, these results appear to be relatively insensitive to the geometric complexity of the coincidence set and free boundary.

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that by \eqref{eq:fe:monotoneproperty} and \eqref{eq:fe:monotoneadditional},
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.  These inclusions are special to multilevel schemes based on defect constraints, and they hold regardless of whether CD operators $\Pi_i$ can be constructed.  A different ``basic'' iteration, an incomplete variant of \pr{cd-mult}, does not need $\Pi_i$ operators.  Application of \pr{icd-tele} requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and \eqref{eq:fe:upwardsuminclusion}.

@@

A convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.  On the other hand, the observation that up-smoothing is more efficient in a multilevel CD context, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.\footnote{Compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009} and \cite{Tai2003}.}

@@

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.  Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the corrected finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\centering
\mbox{\input{tikz/vcycle.tex} \quad
\input{tikz/fcycle.tex}}
\caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets. Right: F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level, and then it applies V-cycles.}
\label{fig:fascdcycles}
\end{figure}

@@

However, an F-cycle (Figure \ref{fig:fascdcycles}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.  (In the VI context one must also truncate to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.

@@

Table \ref{tab:results:classical} shows the resulting iteration counts.  As expected, RS Newton applied by itself as a solver shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.  These experiments over-solve, with \texttt{rtol} $=$ \texttt{atol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.  The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance.  For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution \elb{describe GK results}

\elb{asymp rate figure?}

@@

\begin{example}  \label{ex:results:plap}
We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p=1.5$-Laplacian, a ``fast'' diffusion with diffusivity limit $D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$.  Note that Example \ref{ex:results:sia} applies a doubly-nonlinear version of a $p=4$ ``degenerate'' diffusion operator.  Well-posedness and solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.

This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and the need for robust smoothers when the operator is nonlinear.  Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.  The exact solution is easily computed for any $p$; see the codes accompanying this paper.  For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.  Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions; the result $\tilde w^2$ is visually close to the exact solution.

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands (including imagemagick):
%   cd examples/
%   python3 plap1d.py -fascd_cycle_type multiplicative -mcoarse 4 -levs 3 -figures
%   [save the images into mcd-extended/paper/fixfigs/vcycle/]
%   for X in start2 down2 down1 up0 up1 up2 next2; do mogrify -trim $X.png $X.png; done
\begin{figure}[ht]
\centering
\input{tikz/imagesvcycle.tex}
\caption{Visualization of one FASCD V-cycle iteration for $p$-Laplacian Example \ref{ex:results:plap}, with $J=3$ and a coarsest mesh of $m_0=5$ vertices.  Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines, obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots, and the zero function is dotted.  Vertical scale varies.}
\label{fig:imagesvcycle}
\end{figure}

Iteration results are shown in Table \ref{tab:results:fastplap1d}.  We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).  F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.  If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').

\elb{check if regularization is even necessary when you re-run with \texttt{-fascd\_levels\_snes\_vi\_zero\_tolerance 1.0e-12}}

Note that convergence is at rate $O(h^{2.005})$; compare \cite{ChoeLewis1991}.
\end{example}

% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
\begin{table}[ht]
\centering
\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
\toprule
\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\
\midrule
 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$ \\
\bottomrule
\end{tabular}
\bigskip
\caption{Iterations for V- and F-cycles in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$.}
\label{tab:results:fastplap1d}
\end{table}
